Phylogenetics and the human microbiome.

### 2013-09-24 sbreview.tex
Back to work

\section{Introduction to the human microbiome}

The human microbiome is the collection of microbial organisms that live inside of and on the surface of humans.
In this section I will review what is known about them and their effects.
More bacteria than human cells.

@@

@@

Phylogenetics is primarily used in this realm as a framework with which data can be organized.
It is common for researchers to equate 16S phylogeny with taxonomy.

@@

@@

Non-overlapping reads.
Assemblies.

@@

@@

\cite{hummelen2010deep}
Deep sequencing of the vaginal microbiota of women with HIV

\cite{ravel2011vaginal}
Vaginal microbiome of reproductive-age women

\cite{fredricks2005molecular}
Molecular identification of bacteria associated with bacterial vaginosis

>>>>

\section{Introduction}

Why, as systematic biologists, should we be interested in what these microbial ecologists have done?
System with a lot of data and a lot of questions.
Indeed, just to do ecology you have to do sequencing.
Many of the issues that are raised by high throughput sequencing of beasties that are more commonly of interest for phylogenetics folk are present, and worse, in microbial studies.
Taxonomic issues.

Lack of "obvious" ways of grouping beasties has led to innovation.
It means something that they have developed their own way of using trees, although there is some overlap with previous literature.
Questions of microbial genomics, translated into questions of function, have real consequences.

Why human microbiome versus microbial ecology?
Relevance to health.
We know a lot more about these bugs.
Can manipulate the systems-- feeding studies and gnotobiotic mice.

Opportunities for SB audience to contribute to this field.

\section{The human microbiome}
The human microbiome is the collection of microbial organisms that live inside of and on the surface of humans.
In this section I will review what is known about them and their effects.
More bacteria than human cells.

@@

% \cite{hummelen2010deep} Deep sequencing of the vaginal microbiota of women with HIV
% \cite{ravel2011vaginal} Vaginal microbiome of reproductive-age women
% \cite{fredricks2005molecular} Molecular identification of bacteria associated with bacterial vaginosis

@@

Phylogenetics is primarily used in this realm as a framework with which data can be organized.
It is common for researchers to equate 16S phylogeny with taxonomy.
UniFrac is interesting.

@@

Surprisingly little work on alpha diversity.

@@

Usual challenges.
Non-overlapping reads.
Assemblies.

However, the apparent success of 16s tree based comparisons leads to some interesting questions.

16s tree has taken us remarkably far, but it's clearly limited.
Continued research on the correlation between phylogeny and function.
Can we correct the 16s tree somehow, or what sort of thing can we use that is a better predictor?

There continues to be discussion about whether there is a tree of life.
Examples.
This can be a theoretical one.
What is the correct representation of the genetic ancestry of an organism that allows us to best predict.

Hubbell's neutral theory applied to human microbiome \cite{fierer2012animalcules,costello2012application}.

@@

There are limitations to what we can learn.
Genome doesn't tell all.
Host side is sometimes missing, but experiments are good.

@@
